---
title: "AWSome simulations"
author: "cameron and danny"
date: "07/06/2015"
output: html_document
---

Results for AWS simulations. Ideal plots are based on a static model in
which the price does not fluctuate over time. Simulation results use 
historical spot price information to run monte-carlo simulations to 
estimate how long and how expensive the runs would be if they were
run at that point in time. Models are based on the execution of a few
different processing pipelines:

1. The Configurable Pipeline for the Analysis of Connectomes (C-PAC) pipeline which performs basic structural processing along with functional-connectomics style fMRI processing.
2. The freesurfer pipeline for extracting cortical thickness and other indices of anatomy from structural MRI data.

The pipelines are executed on a SGE style cluster with a single master node and up to 20 compute "worker" nodes. The master node is a t2.small
on-demand instance with local data storage to support the execution of 
all other nodes. 

```{r}
library(ggplot2)
library(reshape)
library(plyr)

if( ! file.exists("cpac_df_agg.csv")){
    cpac_df = read.csv("cpac_df.csv")

    summary(cpac_df)

    # Mean statistics by av_zone, num_datasets, and bid_ratio
    cpac_df_agg<-ddply(cpac_df,.(Av.zone,Num.datasets,Bid.ratio),
                   summarize,
                   mean_total_cost=median(Total.cost),
                   mean_instance_cost=median(Instance.cost),
                   mean_storage_cost=median(Storage.cost),
                   mean_xfer_cost=median(Tranfer.cost),
                   mean_total_time=median(Total.time),
                   mean_run_time=median(Run.time),
                   mean_wait_time=median(Wait.time),
                   mean_upl_time=median(Upload.time),
                   mean_dl_time=median(Download.time),
                   mean_interrupts=median(Interrupts))


    # remove the full df
    rm(cpac_df)

    cpac_df_agg$Region[grep("us-west",cpac_df_agg$Av.zone)]="US West"
    cpac_df_agg$Region[grep("us-east",cpac_df_agg$Av.zone)]="US East"
    cpac_df_agg$Region[grep("ap",cpac_df_agg$Av.zone)]="Asia Pacific"
    cpac_df_agg$Region[grep("eu",cpac_df_agg$Av.zone)]="Europe"
    cpac_df_agg$Region[grep("sa",cpac_df_agg$Av.zone)]="South America"
    cpac_df_agg$Region=factor(cpac_df_agg$Region)
    #cpac_df_agg=subset(cpac_df_agg,Num.datasets<=10000)
    cpac_df_agg=subset(cpac_df_agg,Num.datasets < 3000)
    write.csv(cpac_df_agg,file="cpac_df_agg.csv")
} else {
  cpac_df_agg=read.csv("cpac_df_agg.csv")
  cpac_df_agg=subset(cpac_df_agg,Num.datasets < 3000)
}


```

You can also embed plots, for example:

```{r plots, fig.width=8, fig.height=12}
library(gridExtra)

p1<-ggplot(subset(cpac_df_agg,Num.datasets==1000), 
           aes(x=Bid.ratio,y=mean_total_cost,col=Av.zone))+
      geom_line()+
      facet_grid(Region~.,scales="free_y")+
      theme_bw()+
      theme(legend.position="None",
            axis.title.x = element_text(size=10,colour="black", vjust=-.8),
            axis.title.y = element_text(size=10,colour="black"),
            axis.text.x = element_text(size=8,colour="black", angle=35),
            axis.text.y = element_text(size=8,colour="black"))


p2<-ggplot(subset(cpac_df_agg,Num.datasets==1000), 
           aes(x=Bid.ratio,y=mean_total_time/3600,col=Av.zone))+
      geom_line()+
      facet_grid(Region~.,scales="free_y")+
      theme(legend.position="None")

p3<-ggplot(subset(cpac_df_agg,Bid.ratio==2.5), 
           aes(x=Num.datasets,y=mean_total_cost,col=Av.zone))+
      geom_line()+
      facet_grid(Region~.,scales="free_y")+
      theme(legend.position="None")

p4<-ggplot(subset(cpac_df_agg,Bid.ratio==2.5), 
           aes(x=Num.datasets,y=mean_total_time/3600,col=Av.zone))+
      geom_line()+
      facet_grid(Region~.,scales="free_y")+
      theme(legend.position="None")

p5<-ggplot(subset(cpac_df_agg,Bid.ratio==2.5), 
           aes(x=Num.datasets,y=mean_interrupts,col=Av.zone))+
      geom_line()+
      facet_grid(Region~.,scales="free_y")+
      theme(legend.position="None")

grid.newpage()
layout=grid.layout(2,2)
pushViewport(viewport(layout=layout))

print(p1,vp=viewport(layout.pos.row=1, layout.pos.col=1))
print(p2,vp=viewport(layout.pos.row=2, layout.pos.col=1))
print(p3,vp=viewport(layout.pos.row=1, layout.pos.col=2))
print(p4,vp=viewport(layout.pos.row=2, layout.pos.col=2))
```


Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
